---
title: "exploreCEHR: overview of EHR-oriented transformer from Pang et al. 2024 arXiv:2402.04400"
shorttitle: "autoregressive transformer for EHRs"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{exploreCEHR: overview of EHR-oriented transformer from Pang et al. 2024 arXiv:2402.04400}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction 

The purpose of this R package is to survey some of the resources
used and produced in the operation of an autoregressive transformer
system for predicting health events based on electronic health records.

A dialogue with Claude Sonnet 4.6 extended led to isolation of
a data simulation process and some exploratory training.  A key
objective is estimation of resource requirements for creating
similar models *de novo*.

The flavor of the information to be processed is sketched here.

A "vocabulary" is defined.

```{r lkvoc, message=FALSE}
library(exploreCEHR)
browse_notation()
```

Patient's event histories are tabulated.  This table is limited to
5000 records; `r nrow(arrow::read_parquet(events_path()))` simulated records are available.

```{r lkev}
browse_events()
```
