---
title: "exploreCEHR: overview of EHR-oriented transformer from Pang et al. 2024 arXiv:2402.04400"
shorttitle: "autoregressive transformer for EHRs"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{exploreCEHR: overview of EHR-oriented transformer from Pang et al. 2024 arXiv:2402.04400}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Introduction 

The purpose of this R package is to survey some of the resources
used and produced in the operation of an autoregressive transformer
system for predicting health events based on electronic health records.

A dialogue with Claude Sonnet 4.6 extended led to isolation of
a data simulation process and some exploratory training.  A key
objective is estimation of resource requirements for creating
similar models *de novo*.

The flavor of the information to be processed is sketched here.

# Vocabulary for health event series

A "vocabulary" is defined.  Some semantics of annotations can
be gleaned from comments in the simulation program:

```
simulate_cehr_gpt.py
====================
Generate 100 K realistic longitudinal health event trajectories in the
CEHR-GPT token-sequence format.

Token-sequence schema (per patient)
-------------------------------------
  START_YEAR:YYYY  AGE:XX  GENDER:M/F  RACE:<race>
  [VS] [VT:OP|IP|ED]  concept…  [VE]
  [ATT:D<n>]      <- inter-visit gap (n = 1-1080 days)
  [ATT:LT]        <- gap > 1080 days
  ...
  [EOS]

Inpatient visits additionally contain intra-visit day tokens [IATT:D<n>]
and end with a discharge-facility token [DC:<facility>].

Outputs  (./output/)
---------------------
  sequences.parquet  -- one row per patient; full token sequence as string
  events.parquet     -- MEDS-style long table (subject_id | time | code | domain)
  summary_stats.json -- prevalence, visit counts, sequence-length stats
```

`browse_notation` provides additional details:

```{r lkvoc, message=FALSE}
library(exploreCEHR)
browse_notation()
```

# Elementary tabulation of event series

Patient's event histories are tabulated.  The table in this vignette is limited to
5000 records; `r nrow(arrow::read_parquet(events_path()))` simulated records are available.

```{r lkev}
browse_events()
```

# Data structure for modeling

Event sequences are reformatted to per-patient records.

In the following we break up the string for one patient's
visits to isolate a single visit and the prescriptions
and conditions noted.

```{r lkseqd}
example(view_sequences)
```
