---
title: "Resource use in training"
author: "Vincent J. Carey, stvjc at channing.harvard.edu"
date: "`r format(Sys.time(), '%B %d, %Y')`"
vignette: >
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteIndexEntry{Resource use in training}
  %\VignetteEncoding{UTF-8}
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
---

# Simple measurement

The `nvidia-smi` utility can record aspects of GPU utilization.
We have used

```
nvidia-smi --query-gpu=timestamp,\
utilization.gpu,\
utilization.memory,\
memory.used,memory.free,memory.total,\
power.draw,temperature.gpu --format=csv --loop=1 > gpu_trace.csv &
```

to collect data from an example run.  One kills the associated
process when the run completes.

The data from a test run with 5000 patients (average 1600 events
per patient, model has 4 heads, embedding dimension 128) are plotted in an elementary way:

```{r lkpl, message=FALSE}
par(mar=c(4,3,2,2))
library(exploreCEHR)
example(plot_trace, ask=FALSE)
```

# Measurements from a partial parameter sweep

An illustration of usage as we explore different
numbers of heads and embedding sizes has been
obtained.  Here the number of patients is 10000 with mean number of events per patient 1800.

GPU memory consumption:

```{r dospecial, message=FALSE}
library(ggplot2)
source(system.file("performance", "merge.R", package="exploreCEHR"))
vramplot
```

GPU fractional utilization:
```{r lkfrac}
utilplot
```
